\LARGE \bf Modeling and Control of a Flexible Catheter for Cardiac Ablation

### 2018-05-03 manuscript/manuscript.tex
Modifications to notebooks and readme files
Stopped tracking ignored files

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
The control of human-operated flexible systems is a problem that has generated significant research. A variety of methods have been proposed to eliminate disturbances to flexible systems. By simplifying disturbances by representing them as impulses that generate initial conditions, some researchers have used impulse-based command generation to reduce existing vibration. This approach closely resembles the common open-loop input shaping method. However, existing literature does not thoroughly account for robustness requirements. This paper expands on the concept of initial condition input shaping and proposes a method of providing robustness to variations in natural frequency. The concept of Specified Insensitivity is used to generate these robust input shapers. Experimental results are provided which support these findings.
\end{abstract}

@@

\section{Modeling}
\label{sec:modeling}

@@

\section{Control}
\label{sec:Control}

@@

The resulting equations of motion were created by Kane's method using the Python SymPy module. When linearized about $Q=\dot{Q}=0$, the state-space representation of the dynamic system is given by:
%
\begin{equation}
\label{eq:eqmotion}
\begin{bmatrix}
 \dot{Q}\\
 \ddot{Q}
\end{bmatrix}
=
A
\begin{bmatrix}
 Q\\
 \dot{Q}
\end{bmatrix}
+
B f_{in}
+
\Omega \Gamma

\end{equation} 
%
where $A$ and $B$ are the linearized state transition and input matrices, and $\Omega$ is the disturbance input matrix.

@@

The control objective for this initial study is to hold the catheter endpoint steady at a specified point in $\bar{X}$. Due the the external disturbance induced by blood flow, a model predictive control (MPC) approach will be used to maintain accurate positioning. The general formulation for MPC of a linear system is realized by discretizing the system:
%
\begin{equation}
\begin{align}
\label{eq:discrete-ss}
x_{n+k} &= A x_n + B u_k + \Omega \Gamma_k\\
y_n &= C x_k,
\end{align}
\end{equation}
%
and determining the correct input, $u_k$ which minimizes a quadratic cost function:
%
\begin{equation}

>>>>

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
Atrial fibrillation is a condition in which the cardiovascular substrate exhibits irregular elctrophysiological activity, causing improper firing of the cardiovascular tissue. To ease symptoms from this condition, catheter ablation is increasingly used to isolate regions from which these irregular signals originate. The current state-of-the-art in catheter ablation requires a clinician to manually control the positioning of a flexible catheter within the heart chambers in the presence of significant disturbances due to the motion of the cardiovascular tissue as well as hemodynamic forces. As a result, position control of the catheter tip is nearly impossible. In order to improve the precision of such interventions, better control methods are needed. This work proposes a model predictive control approach which can be used to augment the control imparted by the clinician to stabilize the catheter in the presence of the aforementioned disturbances. 
\end{abstract}

@@

\subsection{Modeling}
\label{sec:modeling}

@@

\subsection{Control}
\label{sec:Control}

@@

The resulting equations of motion were created by Kane's method using the Python SymPy module. When linearized about $Q=\dot{Q}=0$, the state-space representation of the dynamic system is given by:
%
\begin{equation}
\label{eq:eqmotion}
\begin{bmatrix}
 \dot{Q}\\
 \ddot{Q}
\end{bmatrix}
=
A
\begin{bmatrix}
 Q\\
 \dot{Q}
\end{bmatrix}
+
B f_{in}
+
\Omega \Gamma
\end{equation} 
%
where $A$ and $B$ are the linearized state transition and input matrices, and $\Omega$ is the disturbance input matrix.

@@

The control objective for this initial study is to hold the catheter endpoint steady at a specified point in $\bar{X}$. Due the the external disturbance induced by blood flow, a model predictive control (MPC) approach will be used to maintain accurate positioning. The general formulation for MPC of a linear system is realized by discretizing the system:

\begin{equation}
x_{k+1} = A x_k + B u_k + \Omega \Gamma_k
\label{eq:discrete_ss}
\end{equation}

and determining the correct input, $u_k$ which minimizes a quadratic cost function:

\begin{equation}
\label{eq:costfunction}
J = \sum_{i=k}^{k+N} Q (y_d - y_i)^2 + R (u_i - u_{i-1})^2,
\end{equation}

where $N$ is the prediction horizon. In this cost function, the predicted error in the outputs, $y_i$, are evaluated relative to desired values, $y_d$. These values are weighted by the state weights, $Q$. Similarly, the change in controller effort from the previous value is penalized and scaled by the input weight, $R$. The input cost is formulated in this way because the actuator effort is not expected to go to zero in the presence of disturbances and desired displacements. Rather, to follow a smooth trajectory, the change in input from one time increment to another should be minimal.

The output of the catheter model which should be controlled is the position of its tip. Using the model presented in this work, the position is given by:

\begin{equation}
\label{eq:y-output}
y = \sum_{i=0}^{n} l_i \sin{\theta_i},
\end{equation}

and the derivative of the position is:

\begin{equation}
\label{eq:y-dot-output}
\dot{y} = \sum_{i=0}^n l_i \dot{\theta}_i \cos{\theta_i}. 
\end{equation}

It is further assumed that the exogeneous input from the hemodynamic forces can be accurately predicted and incorporated into the MPC framework. Therefore, \eqref{eq:discrete_ss} is used as a constraint for the controller. In addition, the actuator effort is limited below the maximum available level. In total, the optimization problem for the model predictive controller takes the form:

\begin{equation}
\begin{aligned}
\label{eq:posconstraints}
& \underset{t}{\text{minimize}}
& &  \sum_{i=k}^{k+N} Q (y_d - y_i)^2 + R (u_i - u_{i-1})^2\\
& \text{subject to}
& & x_{k+1} = A x_k + B u_k + \Omega \Gamma_k,\\
&&&  u_k \leq u_{max}.\\
\end{aligned}
\end{equation}
